### 2025-01-11 supplement.tex
moved some floats for better whitespace usage
some supplement cleanup
updated plots in the supplement

\input{cakes_shared}
\begin{document}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Supplementary Materials: An Example Article},
  pdfauthor={D. Doe, P. T. Frank, and J. E. Smith}
}
\fi

@@

@@

\begin{figure}
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.95\textwidth]{plots/fashion-mnist-indexing.png}\\
        \subcaption{Fashion-mnist}
        \label{fig:supplement:fashion-mnist-indexing}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.95\textwidth]{plots/glove-25-indexing.png}\\
        \subcaption{Glove-25}
        \label{fig:supplement:glove-25-indexing}
    \end{subfigure}
    \vspace{1em}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.95\textwidth]{plots/sift-indexing.png}\\
        \subcaption{Sift}
        \label{fig:supplement:sift-indexing}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.95\textwidth]{plots/random-indexing.png}\\
        \subcaption{Random}
        \label{fig:supplement:random-indexing}
    \end{subfigure}%
    \\
    \vspace{1em}
    \caption{Indexing and tuning time for each algorithm with each of the ANN benchmark datasets and the Random dataset.}
    \label{fig:supplement:indexing}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=3.4in]{plots/fashion-mnist-knn-100.png}
    \caption{
        Scaling behavior of algorithms on fashion-mnist with $k=100$. 
    }
    \label{fig:supplement:fashion-mnist-k-100}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=3.4in]{plots/sift-knn-100.png}
    \caption{
        Scaling behavior of algorithms on sift with $k=100$. 
    }
    \label{fig:supplement:sift-k-100}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=3.4in]{plots/glove-25-knn-100.png}
    \caption{
        Scaling behavior of algorithms on glove-25 with $k=100$. 
    }
    \label{fig:supplement:glove-25-k-100}
\end{figure}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=3.4in]{plots/random-knn-100.png}
    \caption{
        Scaling behavior of algorithms on random with $k=100$. 
    }
    \label{fig:supplement:random-k-100}
\end{figure}

@@

\begin{figure}[ht!]
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/radius/fashion-mnist-60000.png}\\
    \subcaption{Fashion-mnist}
    \label{fig:results:fashion-mnist-radius}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/radius/glove-25-1183514.png}\\
    \subcaption{Glove-25}
    \label{fig:results:glove-25-radius}
    \end{subfigure}
    \vspace{1em}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/radius/sift-1000000.png}\\
    \subcaption{Sift}
    \label{fig:results:sift-radius}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/radius/radio-ml-97920.png}\\
    \subcaption{RadioML}
    \label{fig:results:radioml-radius}
    \end{subfigure}%
    \\
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/radius/silva-2224640.png}\\
    \subcaption{Silva 18S}
    \label{fig:results:silva-radius}
    \end{subfigure}%  
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/radius/random-1000000.png}\\
    \subcaption{A random dataset}
    \label{fig:results:random-radius}
    \end{subfigure}
    \vspace{1em}
    \caption{Radius vs. cluster depth across six datasets, grouped by decile of radius and weighted by the cardinalities of the clusters.
    The last dataset is randomly generated.}
    \label{fig:results:radius-plots}
\end{figure}

\begin{figure}[ht!]
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/fractal_density/fashion-mnist-60000.png}\\
    \subcaption{Fashion-mnist}
    \label{fig:results:fashion-mnist-fractal_density}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/fractal_density/glove-25-1183514.png}\\
    \subcaption{Glove-25}
    \label{fig:results:glove-25-fractal_density}
    \end{subfigure}
    \vspace{1em}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/fractal_density/sift-1000000.png}\\
    \subcaption{Sift}
    \label{fig:results:sift-fractal_density}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/fractal_density/radio-ml-97920.png}\\
    \subcaption{RadioML}
    \label{fig:results:radioml-fractal_density}
    \end{subfigure}%
    \\
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/fractal_density/silva-2224640.png}\\
    \subcaption{Silva 18S}
    \label{fig:results:silva-fractal_density}
    \end{subfigure}%  
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/fractal_density/random-1000000.png}\\
    \subcaption{A random dataset}
    \label{fig:results:random-fractal_density}
    \end{subfigure}
    \vspace{1em}
    \caption{Fractal Density vs. cluster depth across six datasets, grouped by decile of fractal density and weighted by the cardinalities of the clusters.
    The last dataset is randomly generated.
    Fractal Density is defined as $\frac{cardinality}{radius^{LFD}}$}
    \label{fig:results:fractal_density-plots}
\end{figure}

>>>>

\input{cakes_shared}
\begin{document}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Supplementary Materials: Let them have CAKES: A Cutting-Edge Algorithm for Scalable, Efficient, and Exact Search on Big DataAn Example Article},
  pdfauthor={M. E. Prior, T. J. Howard III, O. McLaughlin, T. Ferguson, N. Ishaq, N. M. Daniels}
}
\fi

@@

\section{Supplementary Methods}

\subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}

We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children that might contain points in the query ball.

To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:supplement:overlapping-children}.
Here, we overload the notation for $\overline{x y}$ to refer both to the line segment joining points $x$ and $y$ as well as to the length of that line segment.

Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively.
Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
Now let $q'$ be the projection of $q$ onto $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
As a consequence of how we assign a point in the parent cluster to the left child in the Partition algorithm, if $\rho < d$, then the left child cannot contain points inside the query ball.
In such a case we proceed to search only the right child.
Otherwise, we proceed with both children.

To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:supplement:overlapping-children}.
By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.

Thus, $d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$. Note, in particular, that this only requires distances between actual points from the dataset, and so it can be used with any distance function, even when $q'$ and $m$ are not actual points or cannot be imputed from the data.

To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, to find the leaf clusters that overlap with the query ball or any clusters which lie entirely within the query ball.
Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, to find all points that are no more than a distance $\rho$ from the query.
The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:supplement:rnn-search-complexity}.

\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \
        \underbrace{
            \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
        }_{\textrm{leaf-search}}
    \Bigg)
    \label{eq:supplement:rnn-search-complexity}
\end{gather}
where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.

\begin{figure}
    \centering
    \includegraphics[scale=0.75]{images/geometry/overlapping-children-3.pdf}
    \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children. Here, $l$ is the left pole, $r$ is the right pole, and $q$ is the query. Other points and distances are described in the text.}
    \label{fig:supplement:overlapping-children}
    \caption{CAKES uses geometric properties of clusters.}
\end{figure}

@@

\begin{figure}
    \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/indexing/fashion-mnist-indexing.png}\\
        \subcaption{Fashion-mnist}
        \label{fig:supplement:fashion-mnist-indexing}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/indexing/glove-25-indexing.png}\\
        \subcaption{Glove-25}
        \label{fig:supplement:glove-25-indexing}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/indexing/sift-indexing.png}\\
        \subcaption{Sift}
        \label{fig:supplement:sift-indexing}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/indexing/random-indexing.png}\\
        \subcaption{Random}
        \label{fig:supplement:random-indexing}
    \end{subfigure}%
    \\
    \caption{Indexing and tuning time for each algorithm with each of the ANN benchmark datasets and the Random dataset.}
    \label{fig:supplement:indexing}
\end{figure}

\begin{figure}
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=1.0\textwidth]{plots/fashion-mnist_PermutedBall_100_throughput.png}
        \subcaption{Fashion-Mnist for $k=100$.}
        \label{fig:supplement:fashion-mnist-scaling}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=1.0\textwidth]{plots/glove-25_PermutedBall_100_throughput.png}
        \subcaption{Glove-25 for $k=100$.}
        \label{fig:supplement:glove-25-scaling}
    \end{subfigure}%
    \\
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=1.0\textwidth]{plots/sift_PermutedBall_100_throughput.png}
        \subcaption{Sift for $k=100$.}
        \label{fig:supplement:sift-scaling}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=1.0\textwidth]{plots/random_PermutedBall_100_throughput.png}
        \subcaption{A random dataset for $k=100$.}
        \label{fig:supplement:random-scaling}
    \end{subfigure}%
    \\
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=1.0\textwidth]{plots/silva-SSU-Ref_PermutedBall_100_throughput.png}
        \subcaption{Silva for $k=100$.}
        \label{fig:supplement:silva-scaling}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=1.0\textwidth]{plots/radio-ml_Ball_100_throughput.png}
        \subcaption{RadioML for $k=100$ at SnR = 10dB.}
        \label{fig:supplement:radioml-scaling}
    \end{subfigure}%
    \\
    \begin{subfigure}[b]{0.94\textwidth}
        \centering
        \includegraphics[width=0.7\textwidth]{plots/legend.png}
        \label{fig:supplement:scaling-legend}
    \end{subfigure}%
    \caption{Throughput (with $k$=100) across six datasets, including a randomly-generated dataset.
    In each plot, the horizontal axis represents increasing cardinality of the dataset, while the vertical axis represents the throughput in queries per second (higher is better).
    For linear search with CAKES, we only report the throughput for a few of the initial multipliers because the trend is clear.}
    \label{fig:supplement:scaling-plots}
\end{figure}

@@

\begin{figure}
    \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/radius/fashion-mnist.png}\\
        \subcaption{Fashion-mnist}
        \label{fig:supplement:fashion-mnist-radius}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/radius/glove-25.png}\\
        \subcaption{Glove-25}
        \label{fig:supplement:glove-25-radius}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/radius/sift.png}\\
        \subcaption{Sift}
        \label{fig:supplement:sift-radius}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/radius/random.png}\\
        \subcaption{A random dataset}
        \label{fig:supplement:random-radius}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/radius/silva-SSU-Ref.png}\\
        \subcaption{Silva 18S}
        \label{fig:supplement:silva-radius}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/radius/radio-ml.png}\\
        \subcaption{RadioML}
        \label{fig:supplement:radioml-radius}
    \end{subfigure}%
    \\
    \vskip 0.005in
    \begin{subfigure}[b]{0.94\textwidth}
        \centering
        \includegraphics[width=0.7\textwidth]{images/radius/legend.png}
        \label{fig:supplement:radius-legend}
    \end{subfigure}%
    \caption{Radius vs. cluster depth across six datasets, grouped by percentile of radius and weighted by the cardinalities of the clusters.
    In order to use the same y-axis for all plots, we divided the radii of all clusters by the maximum radius of any cluster in the dataset.
    Note that for the Silva 18S and the RadioML datasets, we use a logarithmic scale for the y-axis.}
    \label{fig:supplement:radius-plots}
\end{figure}

\begin{figure}
    \captionsetup[subfigure]{aboveskip=-15pt,belowskip=-3pt}
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/fractal_density/fashion-mnist.png}\\
        \subcaption{Fashion-mnist}
        \label{fig:supplement:fashion-mnist-fractal_density}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
            \includegraphics[width=0.9\textwidth]{images/fractal_density/glove-25.png}\\
            \subcaption{Glove-25}
            \label{fig:supplement:glove-25-fractal_density}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/fractal_density/sift.png}\\
        \subcaption{Sift}
        \label{fig:supplement:sift-fractal_density}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/fractal_density/random.png}\\
        \subcaption{A random dataset}
        \label{fig:supplement:random-fractal_density}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/fractal_density/silva-SSU-Ref.png}\\
        \subcaption{Silva 18S}
        \label{fig:supplement:silva-fractal_density}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=0.9\textwidth]{images/fractal_density/radio-ml.png}\\
        \subcaption{RadioML}
        \label{fig:supplement:radioml-fractal_density}
    \end{subfigure}%
    \\
    \vskip 0.005in
    \begin{subfigure}[b]{0.94\textwidth}
        \centering
        \includegraphics[width=0.7\textwidth]{images/fractal_density/legend.png}
        \label{fig:supplement:fractal_density-legend}
    \end{subfigure}%
    \caption{Fractal Density vs. cluster depth across six datasets, grouped by percentile of fractal density and weighted by the cardinalities of the clusters.
    Fractal Density is defined as $\frac{cardinality}{radius^{LFD}}$.
    As with the radius plots, we normalized the cluster radii before calculating the Fractal Density.
    We also excluded all clusters with normalized radii smaller than $10^{-6}$, as they distort the scale of the plots.}
    \label{fig:supplement:fractal_density-plots}
\end{figure}
